\TheTitle

### 2017-07-07 paper/design.tex
Splitting file into subfiles
vim is dying trying to syntax-highlight the lineval.tex file. It's
taking ~2 minutes to move the cursor one line (both in the tty and in
X).

% vim:set et fenc=utf-8 ft=tex sw=2 ts=2 tw=72:

\newif\ifdraft
%\drafttrue

\ifdraft
        \documentclass[draft]{IEEEtran}
        \def\baselinestretch{1}
        % \setlength{\marginparwidth}{1.5cm}
\else
        \documentclass[conference]{IEEEtran}
\fi

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[hyphens]{url}
\urlstyle{same}

%% Citations
\usepackage[nospace]{cite}

%% Table Support
\usepackage{array}
\usepackage{dcolumn}
\usepackage{longtable}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{tabulary}

%% Extra support
\usepackage{xspace}
\usepackage{amsmath}
\usepackage{balance}
\usepackage{placeins}

%% Algorithms
\usepackage{algorithm}
\usepackage{algpseudocode}

%% Graphics
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{pgfplotstable}
\usepackage{xcolor}
\usepackage{color}
\usepackage{listings}

\usetikzlibrary{positioning, arrows}

\input{color.tex}
\input{macros.tex}

\newcommand{\TheTitle}{Merge-Tree: Visualizing the Integration of Commits into Linux}
\newcommand{\TheAuthors}{Evan Wilde, Daniel M. German}
\newcommand{\TheEmails}{etcwilde@uvic.ca, dmg@uvic.ca}
\newcommand{\TheSubject}{Understanding a large number of commits and merges}
\newcommand{\TheKeywords}{Linux, git, visualizations}

%% Referencing
\usepackage[unicode=true,
          pdftitle={\TheTitle},
          pdfauthor={\TheAuthors},
          colorlinks=false]{hyperref}

\algdef{SE}[DOWHILE]{Do}{doWhile}{\algorithmicdo}[1]{\algorithmicwhile\ #1}%

\lstset{frame=tb,
  language=python,
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  columns=flexible,
  basicstyle={\small\ttfamily},
  numbers=none,
  numberstyle=\tiny\color{gray},
  keywordstyle=\color{chartblue},
  commentstyle=\color{chartred},
  stringstyle=\color{chartgreen},
  breaklines=true,
  breakatwhitespace=true,
  tabsize=3
}

\synctex=1

\begin{document}

\title{\TheTitle}

\author{
        \IEEEauthorblockA{\TheAuthors}
        \IEEEauthorblockN{Department of Computer Science,
                University of Victoria, Canada.}
        \IEEEauthorblockA{Email: \TheEmails}
}
\maketitle

\begin{abstract}

  With an average of more than 900 top-level merges into the Linux
  kernel per release, many containing hundreds of commits and some
  containing thousands, maintenance of older versions of the kernel
  becomes nearly impossible.  Various commercial products, such as the
  Android platform, run older versions of the kernel. Due to security,
  performance, and changing hardware needs, maintainers must understand
  what changes (commits) are added to the current version of the kernel
  since the last time they inspected it in order to make the necessary
  patches.

  Current tools provide information about repositories through the
  directed acyclic graph (DAG) of the repository, which is helpful for
  smaller projects. However, with the scale and number of branches in
  the kernel the DAG becomes overwhelming very quickly. Furthermore, the
  DAG contains every ancestor of every commit, while maintainers are
  more interested in how and when a commit arrives to the official Linux
  repository.

  In this paper, we propose the merge-tree, a simplified transformation
  of the DAG of the Linux git repository that shows the way in which
  commits are merged into the master branch of Linux. Using the
  merge-tree, we build \tool, a tool that is designed to allow users to
  explore how commits are merged into the Linux kernel.

\end{abstract}

\begin{IEEEkeywords}
  \TheKeywords
\end{IEEEkeywords}

\section{Introduction}

Between 50k and 70k commits are added to the Linux kernel per version
requiring maintainers of older versions of the kernel to sift through
thousands of commits and merges with tools that are unable to filter and
effectively visualize projects at the scale of the kernel. Older
versions of the kernel are used in embedded systems and mobile phones;
for security purposes, performance needs, and changing hardware
requirements, maintainers must be able to understand the changes being
made in the current version of the kernel in order to produce the
necessary patches for the older versions of the kernel. Tools like Gitk
use a directed acyclic graph (DAG) model of the repository, showing all
commits and merges in chronological order by when the commit was
authored, not by when it arrived in the official Linux repository.

\begin{figure*}
        \centering
        \includegraphics[width=0.97\textwidth]{figures/gitk.png}
        \caption{A view of the Gitk interface centered on merge
          b34870fc9ff15fe46c4066faeeec437a4e63e2d8 by Miller. Commits point toward
          their ancestors and there is no clear path from the commit to the merge
          with the master branch. Neither Gitk nor Git are capable of showing the
          commits in master.}
        \label{fig:gitk}
%\vspace{-4mm}
\end{figure*}

\begin{figure}
        \centering
        \includegraphics[width=0.45\textwidth]{figures/github_viewer.png}
        \caption{Github Failure Message showing the DAG of the Linux Project and its
                relationship to other forks.}
        \label{fig:gitfail}
%\vspace{-2mm}
\end{figure}

This representation works in smaller projects; it enables users to see
when changes are made, when these changes are merged, how each branch is
interacting, and the point where a branch forks from the master branch.
In large modular projects, like the Linux kernel, the DAG becomes a mess
of merges and commits (Figure~\ref{fig:gitk}) losing its visual meaning.
In some cases, the Linux kernel is simply too large for the system to
generate a visualization; Github provides a DAG view for many projects,
but is unable to display the visualization for projects at the scale of
the Linux of the kernel (Figure~\ref{fig:gitfail}).  Between 60k and 70k
new commits are created for the Linux project every year; according to
our previous work\cite{German2015}, a commit takes a median of 30 days
from the time it is authored until it arrives in the official
repository. The snapshot of the kernel tomorrow may be different than
the snapshot from today, containing new commits authored in the past;
distinguishing these new commits from the commits in the snapshot from
today is not trivial.

One major challenge with visualizing the arrival of commits to a
repository is that Git does not store the date that a commit was merged
into another branch, including the master branch. To complicate the
problem, the DAG only has references to the ancestors of a commit (a
model necessary for the operation of Git), but maintainers would prefer
knowing the path a commit followed to reach the master repository.
Tracing a path that any commit followed to the master repository would
imply that for any given merge, it would be possible to know which
commits were merged. A user could inspect the commits that arrived into
the master branch within a given time-frame by checking which commits
were merged during that time-frame.

This paper makes two contributions; first, we describe a method of
converting the DAG of the Linux repository into a tree, or
\emph{merge-tree} of the repository, that represents the path used by a
commit to reach the master branch; second, we present a method to
inspect and visualize the history of merges in the Linux project using
the merge-tree model.

These methods and visualizations are implemented in a web-based tool
called \tool\footnote{\tool is currently available at
  \url{http://li.turingmachine.org}}.  Our visualizations and tool
provide information about the location of any given commit or merge in
its respective merge-tree, the files edited, the modules edited, and the
commit message. \tool allows users to apply various filters, including
the release version, along with a keyword or phrase from the log
preview, the name of the author, or the commit ID. The user can request
all merges made by Linus that contain a commit or inner merge that
matches the search query, or just the commits and merges that match the
query.

\section{Merge-Tree model}
\label{sec:mergetree}

Git uses a directed acyclic graph (DAG) as its main data model. In this
model, a commit has one or more ordered parents (ancestors), except the
root commits of a repository which do not have ancestors--Linux's git
repository has two such commits.  Commits are divided into merge
commits, merging two or more branches, and non-merge commits. Non-merge
commits have only one parent, while merge commits have two or more. The
order of the parents matter: the first parent is the branch in which the
merge is being done, while the rest indicate the branches being merged
(Git allows merging multiple branches simultaneously). The second parent
is the first branch merged, the third parent is the second branched
merged, etc.

Once a commit is created, it is never changed. Git allows operations to
alter commits or reorder them, but it changes the commit ID in the
process, effectively replacing it with a new commit. This makes commits
unable to record the traversal of merges from the commit to the merge
into the master branch.

A short example: assume the commits represented in
Figure~\ref{fig:repoEvents} show the sequence of events in a repository.
In this case, commits are performed in various repositories and
branches. The DAG representation of the commits is shown in
Figure~\ref{fig:repoDAG}. Notice that the DAG loses information about
the master branch and the repository that the master branch is part of.
The merge-tree view of this DAG is visible in Figure~\ref{fig:repoTree}.
Note that the direction of the edges of the DAG have been inverted,
instead of pointing from the child to the ancestors, it points from the
ancestor to its successors, forming a path to the master branch. Also
note that the DAG has been simplified, showing only a single edge on the
path to master for any commit.

\begin{figure}[htbp]
  \centering
  \begin{tikzpicture}[auto, on grid, semithick, state/.style={circle, text=black}]
    \foreach \x in {0, 1, 2, 3, 4, 5, 6, 7}
    \draw[shift={(\x + 0.5, -0.5)}, color=black] (0cm, 4cm) -- (0pt, -0.2cm);

    \node[state, draw=chartblue] (1) {1};
    \node[state, draw=chartyellow, above right= of 1] (2) {2};
    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 2] (3) {3};
    \node[state, draw=chartblue, right= 2cm of 1] (4) {4};
    \node[state, draw=chartyellow, above right=of 4] (5) {5};
    \node[state, draw=chartred, above right=of 5] (6) {6};
    \node[state, draw=chartyellow, right=of 5] (7) {7};
    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 7](8) {8};
    \node[state, draw=chartyellow, right= 2cm of 7] (9) {9};
    \node[state, draw=chartyellow, right=of 9] (10) {10};
    \node[state, draw=chartblue, below right=of 9] (11) {11};
    \node[state, draw=chartblue, below right=of 10] (12) {12};

    \draw (12) edge[-stealth] (11) edge[chartyellow, -stealth] (10);
    \draw (11) edge[-stealth] (4) edge[chartyellow, -stealth] (9);
    \draw (10) edge[chartyellow, -stealth] (9);
    \draw (9) edge[chartmagenta, -stealth] (8) edge[chartred, -stealth] (6)
              edge[chartyellow, -stealth] (7);
    \draw (8) edge[chartmagenta, -stealth] (7);
    \draw (7) edge[chartyellow, -stealth] (5);
    \draw (6) edge[chartred, -stealth] (5);
    \draw (5) edge[chartmagenta, -stealth] (3) edge[chartyellow,-stealth] (2)
              edge[chartyellow, -stealth] (4);
    \draw (4) edge[-stealth] (1);
    \draw (3) edge[chartmagenta, -stealth] (2);
    \draw (2) edge[chartyellow, -stealth] (1);

    \node [draw=chartblue, below = 1.5cm of 1] (l1) {Master};
    \node [draw=chartyellow, right = 1.5cm of l1] (l2) {Repo A};
    \node [draw=chartred, right = 2.5cm of l2] (l3) {Branch of Repo A};
    \node [draw=chartmagenta, right= 2.5cm of l3] (l4) {Repo B};

        \foreach \x in {0, 1, 2, 3, 4, 5, 6, 7, 8}
    \node[shift={(\x, -0.6)}, color=black] {$t_\x$};
  \end{tikzpicture}
  \caption{Example of a sequence of events performed in different
    repositories. The horizontal axis represents time. Each horizontal
    section represents a different branch and/or repository. Each commit
    points to its ancestor.}
  \label{fig:repoEvents}
%\vspace{-3mm}
\end{figure}

\begin{figure}[htbp]
  \centering
  \begin{tikzpicture}[auto, on grid, semithick, state/.style={circle, text=black}]
    \node[state, draw=chartblue] (1) {1};
    \node[state, draw=chartyellow, above right= of 1] (2) {2};
    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 2] (3) {3};
    \node[state, draw=chartblue, right= 2cm of 1] (4) {4};
    \node[state, draw=chartyellow, above right=of 4] (5) {5};
    \node[state, draw=chartred, above right=of 5] (6) {6};
    \node[state, draw=chartyellow, right=of 5] (7) {7};
    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 7](8) {8};
    \node[state, draw=chartyellow, right= 2cm of 7] (9) {9};
    \node[state, draw=chartyellow, right=of 9] (10) {10};
    \node[state, draw=chartblue, below right=of 9] (11) {11};
    \node[state, draw=chartblue, below right=of 10] (12) {12};

    \draw (12) edge[-stealth] (11) edge[chartyellow, -stealth] (10);
    \draw (11) edge[-stealth] (4) edge[chartyellow, -stealth] (9);
    \draw (10) edge[chartyellow, -stealth] (9);
    \draw (9) edge[chartmagenta, -stealth] (8) edge[chartred, -stealth] (6)
              edge[chartyellow, -stealth] (7);
    \draw (8) edge[chartmagenta, -stealth] (7);
    \draw (7) edge[chartyellow, -stealth] (5);
    \draw (6) edge[chartred, -stealth] (5);
    \draw (5) edge[chartmagenta, -stealth] (3) edge[chartyellow,-stealth] (2)
              edge[chartyellow, -stealth] (4);
    \draw (4) edge[-stealth] (1);
    \draw (3) edge[chartmagenta, -stealth] (2);
    \draw (2) edge[chartyellow, -stealth] (1);
  \end{tikzpicture}
  \caption{DAG representation of the commits represented in
    Figure~\ref{fig:repoEvents}. The DAG loses information about which
    repository the commit is performed in and through which merges it
    has passed on its way to the master branch. The DAG does not even
    distinguish the master branch from other branches.}
  \label{fig:repoDAG}
%\vspace{-3mm}
\end{figure}

\begin{figure}[htbp]
  \centering
  \begin{tikzpicture}[auto, on grid, semithick, state/.style={circle, text=black}]

    \draw[chartblue]
      (-0.5, -0.5) -- (8.5, -0.5) -- (8.5, 0.5) -- (-0.5, 0.5) -- (-0.5, -0.5);

    \node[state, draw=chartblue] (1) {1};
    \node[state, draw=chartyellow, above right= of 1] (2) {2};
    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 2] (3) {3};
    \node[state, draw=chartblue, right= 2cm of 1] (4) {4};
    \node[state, draw=chartyellow, above right=of 4] (5) {5};
    \node[state, draw=chartred, above right=of 5] (6) {6};
    \node[state, draw=chartyellow, right=of 5] (7) {7};
    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 7](8) {8};
    \node[state, draw=chartyellow, right= 2cm of 7] (9) {9};
    \node[state, draw=chartyellow, right=of 9] (10) {10};
    \node[state, draw=chartblue, below right=of 9] (11) {11};
    \node[state, draw=chartblue, below right=of 10] (12) {12};

    \draw (2) edge[chartyellow, -stealth] (5)
          (3) edge[chartmagenta, -stealth](5)
          (5) edge[chartyellow, -stealth] (7)
          (7) edge[chartyellow, -stealth] (9)
          (6) edge[chartred, -stealth] (9)
          (8) edge[chartmagenta, -stealth] (9)
          (9) edge[chartyellow, -stealth] (11)
          (10) edge[chartyellow, -stealth] (12);

  \end{tikzpicture}
  \caption{Merge-tree view of the commits  represented in
    Figure~\ref{fig:repoEvents} showing the path they followed to reach
    the master branch. In this model the successors of each commit
    represents the path followed by that commit to reach the master
    branch.}
  \label{fig:repoTree}
\vspace{-3mm}
\end{figure}

\subsection{Computing the merge-tree of the DAG of Linux}

Computing the merge-tree from a DAG for any repository may not be
possible; however, certain features of the development process of Linux
make it feasible to compute the merge-tree for the Linux repository.
First, the master branch of Linux is maintained by Linus Torvalds, and
only Linus has write access to it. We have verified this assertion in
previous research~\cite{German2015}. We have developed a heuristic that
is presented in Algorithm~\ref{fig:alg}. In short, the algorithm first
identifies the commits made directly to the master branch; whereafter it
recursively determines the shortest path (in terms of time), using the
DAG, from each commit to the master branch using the inverted DAG.

\begin{algorithm}
        \caption{Computing the merge-tree of Linux Git's DAG}\label{fig:alg}
        \begin{algorithmic}
                \Function{ComputeMergeTree}{DAG}: tree
                \State {\# Compute the tree from the DAG of Linux repository.}
                \State {\# Returns $Tree$, a graph containing every commit }
                \State {\# in DAG with the path it followed to master.}

                \State $head \gets \textit{Head of master of git repository}$
                \State $master \gets \textit{traverse DAG from head using }$
                \State \quad\quad\quad\quad $\textit{first ancestor until reaching root}$
                \State $nodes(Tree) \gets nodes(DAG)$
                \State \Function{distance2Master}{cid} : seconds
                \State {\# Helper function}
                \State {\# Recursively compute shortest distance to master}
                \State {\# setting cid's successor (next) in its way to master.}
                \State {\# This function should be memoized. Otherwise it}
                \State {\# would run in exponential time.}
                \If {\textit{cid in master}}
                \State \Return 0
                \EndIf
                \State    $d \gets 	\infty$
                \State {\# Traverse the inverted DAG}
                \For{$c \in children(cid, DAG)$}
                \If {$c \in master$}
                \State $d_1 \gets commitTime(c)-commitTime(cid)$
                \Else
                \State {$d_1 \gets distance2Master(c)$}
                \EndIf
                \If {$d_1 < d $}
                \State $next \gets c$
                \State  $d \gets d_1$
                \EndIf
                \EndFor
                \State {\# $c$ is the commit that follows $cid$}
                \State {\# in its way to master}
                \State add edge $(cid, next)$ to $Tree$
                \State \Return $d$
                \EndFunction

                \State {\# Compute the distance for each commit}
                \State {\# discarding result}
                \For{$c \in nodes(DAG)$}
                \State $distance2Master(c)$
                \EndFor
                \State \Return $Tree$
                \EndFunction
        \end{algorithmic}
\end{algorithm}

\subsection{Evaluation}

Merges that do not have conflicts provide information to verify this
heuristic. If a merge does not contain a conflict, it records a summary
of the commits that it merges. See Figure~\ref{fig:sampleMerge} for an
example. This summary contains a list of the first 20 non-merge commits
in the merge, including their one-line log description, the full logs of
the merge commits that merge this subset, and the total number of
non-merge commits in the merge.

\begin{figure}[htbp]
        \centering
        {\fontsize{7}{9}
        \begin{verbatim}
Merge: 8cbd84f fd8aa2c
Author: Linus Torvalds <torvalds@linux-foundation.org>
Date:   Tue Aug 10 15:38:19 2010 -0700

Merge branch 'for-linus' of git://neil.brown.name/md

* 'for-linus' of git://neil.brown.name/md: (24 commits)
md: clean up do_md_stop
[... edited for the sake of space]
md: split out md_rdev_init
md: be more careful setting MD_CHANGE_CLEAN
md/raid5: ensure we create a unique name for kmem_cache...
...
        \end{verbatim}}\vspace{-5mm}
        \caption{Example of how merges record a subset of commits being merged. The
                commit only shows the first 20 one-line summaries messages for the 24
                non-merge commits it merged. The ending ``...'' is part of the log
                and represents that other commits were merged.}
        \label{fig:sampleMerge}
\end{figure}

We used this information to evaluate the accuracy of the merge-tree
model extracted from the DAG. The method we followed started with the
extraction of the commit history up to July 20, 2016. We computed the
merge tree of every commit until then. Since Linus Torvalds mostly does
merging directly into master, we assumed that every merge by him is the
root of a merge-tree. As described above, the log of a merge-commit
usually contains the number of commits in the merge the first 20
summaries of commits being merged. We extracted merges by Linus Torvalds
using the command \mycode{log --merges --author='Torvalds} and compared
the number of commits according to the log with the number of commits in
the  merge-tree rooted in this commit. We also used the summaries of the
commits found in the merge (not necessarily all---see above) to make
sure those commits were in their corresponding merge tree. For example,
for the merge in Figure~\ref{fig:sampleMerge} we would expect that the
merge tree rooted at \mycode{8cbd84f} contains 24 commits, and the
one-line summaries corresponds to commits in that merge-tree. We also
inspected those with differences to make sure they were true errors.
The results can be summarized as follows:

\begin{itemize}
\item Five merges were false-errors because their logs did not contain accurate information (were probably edited by
  hand). For example in \mycode{42a579a0f...} one commit summary was missing (the line was empty), in \mycode
  {c55d267} the summaries were reordered.
\item The heuristic correctly identified that 79 of Linus merges (between Jun 7, 2014 and Jun 2, 2014) were made to a
  branch (not master). This branch was merged at \mycode{3f17ea6d..} which contained 6809 commits.
\item The heuristic worked perfectly until Sept 4, 2007, the earliest date that it could be verified.
  Before this date, and until Dec 12, 2006, merges did not include a summary of the commits they included, hence making it
  impossible to verify; during this period, however, we correctly identified the merges by Linus into master.
\item Before Dec. 12, 2006 (1542 merges) our heuristic breaks due to the
presence of a \textit{foxtrot} commit (\mycode{c436688...}), which confounded the true master branch of a
repository (see \url{http://bit-booster.blogspot.ca/2016/02/no-foxtrots-allowed.html} for a description of the issue).
\end{itemize}

In summary, of the merges after Sept 4, 2007, in 100\% of them (16,860)
our heuristic was correct. It failed in 1,542 commits before Dec. 12,
2006 and in 836 it appears to be correct (Dec 7, 2006 to Sept 4, 2007).

\section{Visualizing the merge-tree of Linux}

The goal of \tool is to simplify the navigation of the kernel commit
information, specifically focusing on merges. This is done by leveraging
the merge-tree to inspect how commits are merged on the path to the
master repository.

\subsection{Use cases}

We designed \tool with two use-cases in mind, though a user may switch
between the cases as they work.

\noindent \textbf{Use-case 1: top-to-bottom
  approach}\label{sec:usecase1}\\ These are users that are maintaining a
section of the kernel and would like to pick a merge (including all the
commits that it merges) and merge it directly into their current
repository. This is useful for reducing the amount of re-implementation
work. For these users, it is important to have the ability to aggregate
metadata about files and modules being effected by the merge. Also, it
is important for these users to be able to navigate from the root of the
merge-tree toward the leaves.

\noindent \textbf{Use-case 2: bottom-to-top
  approach}\label{sec:usecase2}\\ These are users that start with a
known merge or commit and would like to see what other changes are being
made in commits that are in the same merge, including knowing the
merge-tree they belong to. This is useful to see what other commits are
related to the current commit and how they get collated into merges that
eventually end in the master branch. This is primarily for maintainers
that need to perform some specific cherry picking of commits. We must
provide these users a mechanism for navigating from a single commit
toward the master branch, allowing them to see other commits that might
be related to their original commit.

\subsection{Data Model}

In our visualizations, we leverage the merge-tree model described in
Section~\ref{sec:mergetree}. In this model each commit is either already
in the master branch, or is part of a tree which is rooted in the merge
that merged it into the master branch.  Each commit, whether a merge or
non-merge, has only one successor; the root of each tree has none as it
was made by Linus Torvalds directly into the master branch. Non-merge
commits contain the metadata for the changes made.  This metadata
includes the files changed, the lines added and removed from each file,
the author, the date the commit was merged into the merge that led to
being merged into the kernel, the date the commit was authored, the
patch, and the commit log. Merges contain less metadata, only storing
the author of the merge, the log, the commit date, and the author date,
and potentially, changes necessary to address conflicts during the
merge. The details of the model are outlined in \cite{German2015}.

\section{Design and Implementation}

@@

\section{Discussion}

From a developers point of view, the are two major disadvantages of the
DAG model of Git: a) its edges point backwards, i.e. commits point to
their ancestor, not to the commit that succeeds them on the path to
being integrated; and b) the DAG contains many more edges that are
necessary to understand how integration occurred. We addressed these two
issues with the creation of the merge-tree model from the DAG of Linux.
Effectively, the merge-tree recovers the details of how each
branch---and the commits they contain---was merged into the master
branch.

In our experience, no other Git repository reaches the level of DAG
complexity that Linux has. In most Git repositories, most merges are not
nested---most merges merge directly into the master repository, and
these merges only merge a few commits. Even in these simplified cases,
the merge-tree can provide a valuable summary of how commits are
integrated into the master branch, specifically since the time of
integration may be very different from the time the commits were
authored.

\evan{Check accuracy of this paragraph}

The biggest challenge is computing the merge-tree of any given
repository. It is likely that our heuristic for computing the merge-tree
will not work with other repositories. The heuristic requires that there
are no fast-forward merges into the master branch of the repository (a
majority of repositories), and that there are no foxtrot merges (a
practice that is starting to be considered desirable by git
users\footnote{See
  \url{http://devblog.nestoria.com/post/98892582763/maintaining-a-consistent-linear-history-for-git}
  and
  \url{https://developer.atlassian.com/blog/2016/04/stop-foxtrots-now/}}).
We are able to use our heuristic with the Linux kernel because of the
strict integration model imposed by Linus. To validate the results of
the heuristic, we use the continuous mining technique described in
\cite{German2015}.  Continuous mining of the repository requires
foresight and planning.

Because \tool leverages the merge-tree model of the kernel, it provides
mechanisms and visualizations that other tools are unable to produce. It
gives us the ability to see how commits are being merged into the master
branch, allowing us to better understand how code is integrated into the
kernel.

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/042dd_DAG.png}
        \caption{Merge Dag View}
        \label{fig:dag_view}
\end{figure}

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/042dd_tree.pdf}
        \caption{Merge tree topology for the same merge as in
                Figure~\ref{fig:dag_view}.}
        \label{fig:tree_view}
\end{figure}

Our goal was to build a tool to enable maintainers to effectively
navigate and browse the changes performed to the kernel over the period
of a release. Achieving this goal includes removing information that
does not pertain to the area of the kernel that the user is interested
in. In Figures~\ref{fig:dag_view} and \ref{fig:tree_view} we can
visually compare the results returned from Gitk and our tool for the
top-level merge ``042dd60ca6dec9a02cefa8edd67de386e35755d6'' from kernel
version 3.10. Note that Gitk has no way to show or recognize that this
is a merge into the master branch.

The log information to this merge-commit in both of these figures is
identical, but the presentation drastically changes our ability to
comprehend what we are seeing.  The primary difference is the removal of
irrelevant information. In Gitk (Figure~\ref{fig:dag_view}), the
visualization results include commits and merges from other components
of the kernel, while our tool (Figure~\ref{fig:tree_view}) only includes
the results that are specific to the component of the kernel that we are
interested in.

Even though this merge is relatively small, containing only four
sub-merges, three of which only contain a single commit, and the last
containing two commits, it is already hard to visualize in Gitk.  With
\tool we are able to immediately see what section of the kernel these
commits and merges pertain to. The DAG view of this provides almost no
explanation, furthermore, users must work to determine where this merge
ends and the next one begins.

\tool further enhances our understanding by summarizing the files and
modules that were edited in the entire branch being merged. We are able
to determine that three files, ``palamas-regulator.c'' had two lines
added and two lines removed, ``dbx500-prcmu.c'' had 12 lines added and
12 lines removed, and ``core.c'' had 5 lines added and 2 lines removed.
Finally, we are able to determine that only the ``regulator'' module was
modified in this merge and was modified by 5 commits.

\section{Related Works}
\label{sec:related_works}

\input{relatedwork.tex}

\section{Controlled User Study}
\label{sec:study}

\dmg{start by describing what you mean by user study... to evaluate linvis we performed a user study...}

\dmg{and you are evaluating the visualizations, NOT the DAG}

We use a controlled mixed-methods users study in order to evaluate the
ability for the DAG to provide users with a conceptual image of how a
commit is integrated into a project, to compare the DAG and \mt models,
and determine which model the users preferred.

In this section, we describe the methods used to for designing and
implementing the user study.
\dmg{no, you describe the study, which includes the methods}

\dmg{Start by giving an overall overview of the study: the study consists of a set of tasks that users are expected to
  perform... do a top down description of what users will do
}

\subsection{Questions and Tasks}
\label{sub:questions}

\dmg{use tasks, not questions. you are asking them to do something that requires an answer, that is different than a question}

\begin{table*}[htpb]
  \centering
  \caption{Conceptual Tasks }
  \label{tab:conceptual_tasks}
  \begin{tabulary}{0.9\textwidth}{LL}
    \toprule
    Task & Description\\
    \midrule
    T1 & Draw a diagram showing how this commit was merged into the master branch, along with any other related commits\\
    T2 & How many individual commits are related to this commit?\\
    T3 & How many merges are involved with merging this commit into the master branch?\\
    \bottomrule
  \end{tabulary}
\end{table*}

\begin{table}[htpb]
  \centering
  \caption{Summarization Tasks}
  \label{tab:summarization_tasks}
  \begin{tabulary}{\linewidth}{LLL}
    \toprule
    Task Set   & Task & Description\\\midrule
    Merge      & T4   & What is the series of merges involved with merging this
    commit?\\
               & T5   & What other commits are merged?\\
    Authorship & T6   & How many authors are involved?\\
               & T7   & Who contributed the most changes?\\
    Files      & T8   & How many files were modified?\\
               & T9   & Which file had the most changes?\\
    Modules    & T10  & Which modules does this merge tree involve?\\
    \bottomrule
  \end{tabulary}
\end{table}

\dmg{avoid editorializing your text: remove purely}

\dmg{watch for tense... we don't begin... the tasks begin... or better, the first tasks...}

The questions and tasks are broken down into four groups. The first two
groups are purely quantitative in nature; the study begins with
conceptual tasks to understand whether the DAG is capable of providing
users with an understanding of how a commit is merged into the master
branch, and other commits that it is merged with. We continue with
summarization tasks, to quantitatively understand if the \mt model makes
a difference in summarization tasks.

The third task set is purely qualitative, \dmg{and are their goal is ... don't imply content in your sentences... be
  precise; this applies to the paragraph before} to gain better understanding
of whether participants preferred working with the \mt model in \tool,
or working with the DAG in Gitk, and what aspects they preferred from
each tool. The final group of questions is simply to gain a better
understanding of the demographic of our participants.

\textbf{Conceptual Tasks}

\dmg{given the choice, not allowed.  I don't understand this section.
  You need to be more top-down in your description. you need to provide
  context on what the users are expected to do, and why. So what are the
  tasks? You never explain.  what is the goal of the tasks? how many?
  what were the tasks? what is expected from each? you need to answer
  these questions, not necessarily in the same order.  also, remember
  this is a comparative study, so I presume the users were asked to do
  both with linvis and without it (choice of gitk/git command line) }

The participants were allowed to use Gitk or the git command line tools
to perform these tasks. These tasks are asked in the specified order,
allowing tabularr to first, create a conceptual understanding of the
commits and merges that will be involved with the rest of the tasks that
follow.tabularmay then use their diagrams to answer T2 and T3. We do not
collect metrics from \tool for the conceptual tasks, opting instead to
determine if the DAG is sufficient for understanding how a commit is
integrated. \tool uses the \mt model show a tree visualization of how
commits are merged, until the commit reaches the master branch.
Table~\ref{tab:conceptual_tasks} shows the conceptual tasks.

\dmg{I don't like the present tense, but that is your option}

We provide the participant with 10 minutes per commit to complete the
first task. The answers to the two other questions in this section are
drawn from the answer in the first task.

\textbf{Summarization Tasks}

\dmg{again, you are going into details first. First the tasks --see
  above-- including their grouping.  then at the end, describe the
  randomization and why it is done (which you did) }

The summarization tasks (Table~\ref{tab:summarization_tasks}) are
grouped into \dmg{how many?} set of tasks according to their theme. The
tasks within a given task set are shuffled, and the task sets are
shuffled, to ensure that similar questions are together, but not in the
same order, and to ensure that the themes are in randomized order
between participants. The order of the commits is shuffled, as well. The
order of the tool is randomized, choosing to start with either \tool or
Gitk, to ensure that we do not place a bias on one tool through the
experiment.

As understanding where the merge begins and ends is vital to answering
the other question, the merge task set will always be the first task set
presented; however, the questions will still be re-ordered within the
task set.

\textbf{User Opinion and Profile}

Finally, we ask the user for their opinions on the tools and for some
information about their level of experience with git.  \dmg{be more
  precise: understanding or preference is too broad. What you want to
  know answers to specific questions regarding how the tools compare and
  which ones they would prefer} The first three questions are intended
to provide us with a better understanding of preference, and insight on
why, the participant had that preference.  The last three questions in
this section are aimed at getting a better understanding of the
demographic of our participants.

\begin{enumerate}
  \item Given these tasks again, which tool would you prefer to use?
  \item Which aspects of each tool did you like and why?
  \item How long have you used git?
  \item If you have used git, for what kind of projects? (personal,
    school courses, professional?)
  \item If you have used git, how many commits, files, and contributors
    were involved with the largest repository you have worked with?
\end{enumerate}

\textbf{Procedure}

\dmg{tell me first why you need to create a script, so I know what you
  do something: we did random assignment of tasks (explain what you did)
  and randomized the order in which the tasks were performed.  Then
  explain why you need a script}

\dmg{it does not randomize the order the tools. It randomizes which tool
  is going to be used on each tasks.  be precise with your language}

\dmg{what you explained how commits relate to tasks and how they get
  assigned? think as if you are explaining this to somebody who has to
  do this: you are the designer, explaining the why and the how }

For each participant, we run a simple python script to generate the
script for the study, eliminating the possibility of bias from us. The
script randomizes the order of the tools, the order of the commits, the
order of the task sets, and the order of the tasks within each task set,
where appropriate.

\subsection{Commit Selection}
\label{sub:commit_selection}

\dmg{see my comments in previous section. it is ok to explain them here,
  but you need to explain what they are briefly in the previous section}

We chose two commits for use in the study. The order that the commits
are presented is randomized between participants, but the order is kept
consistent through the tasks described in the next subsection.

\dmg{this place does not require a description of the database. This
  requires just to say the commits/merge set (dates and numbers are ok)
  only} We use the \tool database, containing commits and merge-tree
information from April 16th 2005 to October 14th, 2014, which
corresponds to being between Linux release 2.6.12-rc3 and Linux
3.17-rc1. \evan{Cite the database even more?} \dmg{again, i have
  problems with your writing: ``we wanted to use in the study commits of
  different size, to determine if the size had an impact in the
  completion of the tasks. We chose commits of three different sizes...}
We chose the commits based on tree sizes. We \dmg{we didn't find: this
  is a fact, not a discovery; what is a majority? 50\%+1? be precise}
\dmg{drop the word inner, and do it in order: 25\% of trees have only
  one commit, 50\%.., 75\%...} found that a majority of the trees
contain at most seven inner commits, while more than 25\% of the trees
contain a single commit. 75\% of the tree contained up to 51 inner
commits and merges, and finally the largest tree contained 7217 nodes.
We chose to work with the trees in \dmg{if you are going to use the
  words quartile, just use them before in your description eg: the
  distribution of the sizes of the trees is x, y merges, whose size
  corresponded to the 25 ,50 quartiles, simple and precisee} the first
and second quartile, as merge trees of sizes between one and seven, not
including the merge into the master branch, make up the majority of the
trees in the database.

\dmg{remove from here, what is here? this is not a road trip :)} From
here, we selected one tree from the trees of a single commit at random.
Selecting a commit from that tree is trivial, as there is only one to
choose.\dmg{remove the obvious, instead concentrate on why the others
  are not trivial}

We selected one tree of size seven at random from all trees of size
seven\dmg{how many where there?}. We placed a restriction the trees, the
selected tree must include at least one internal merge to increase the
complexity of the trees tested. For each randomly selecting a tree, we
chose one of its internal commits at random (using the
\verb|random.choice| function in python 3.6.1).

>>>>

% vim:set et sw=2 ts=4 tw=72:

@@
